---
title: "Nordic Council Sustainable Development Indicators"
author: "Prepared by: Scott I. Large"
date: "17 March 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# devtools::install_github("ICES-dk/rICES", force = FALSE)
# devtools::install_github("hadley/readxl", force = FALSE)
library(rICES)
library(readxl)
library(dplyr)
library(reshape2)
library(RColorBrewer)
library(dplyr)
library(knitr)
library(XML)
library(stringr)
library(ggplot2)
# 
plotDir = "output/"
stockTable <- rICES::getSummaryTable(year = 2015)
# 
```

## Guild and ecoregion

The following stocks, guilds, and ecoregions have been included in the overview. F~MSY~ F, SSB, and MSY B~trigger~ were downloaded from [ICES Stock Assessment Graphs](http://ices.dk/marine-data/tools/Pages/stock-assessment-graphs.aspx). Click the links to see the [aggregated data](https://github.com/ICES-dk/rStockOverview/blob/master/NordicCouncilRefPoints.csv) and [R code](https://github.com/ICES-dk/rStockOverview/blob/master/nordicIndicators.rmd) used to download and plot  [Figure 1](https://github.com/ICES-dk/rStockOverview/blob/master/output/NordicCouncilRefPoints_ALL.png) and [Figure 2](https://github.com/ICES-dk/rStockOverview/blob/master/output/NordicCouncilRefPoints_ecoregion-guild.png).

For questions, please email <A HREF="mailto:scott.large@ices.dk?subject=Nordic Council plots">scott.large@ices.dk</A>.

```{r guildlist, echo=FALSE, include=TRUE}
# 
guildList <- read.csv("~/git/ices-dk/rStockOverview/fisheryGuild.csv")
# 
guildList <- guildList %>%
  rename(Stock = Stock.code,
         GUILD = Fisheries.Guild) %>%
  filter(GUILD %in% c("pelagic", "benthic", "demersal")) %>%
  mutate(Stock = tolower(Stock),
         Stock = gsub("[[:space:]]", "", Stock))
# 

overTable <- guildList %>%
  inner_join(stockTable, c("Stock" = "STOCKID")) %>%
  rename(ECOREGION = EcoRegion) %>%
  distinct(ECOREGION, GUILD, Stock) %>%
  select(ECOREGION, GUILD, Stock, FMSY, MSYBtrigger) %>%
  filter(ECOREGION %in% c("North Sea", "Iceland and East Greenland",
                          "Widely distributed and migratory stocks",
                          "Baltic Sea", "Faroe Plateau Ecosystem",
                          "Barents Sea and Norwegian Sea")) %>%
  filter(!is.na(FMSY) | !is.na(MSYBtrigger)) %>%
  arrange(ECOREGION, GUILD)
# 
knitr::kable(overTable, format="markdown")
```


```{r stocktable, echo = FALSE}
# 
df <- guildList %>%
  inner_join(stockTable, c("Stock" = "STOCKID")) %>%
  rename(ECOREGION = EcoRegion) %>%
  filter(SpeciesName != "Psetta maxima (historic name)" &
           Stock %in% unique(guildList$Stock) &
           ECOREGION %in% c("North Sea", "Iceland and East Greenland",
                          "Widely distributed and migratory stocks",
                          "Baltic Sea", "Faroe Plateau Ecosystem",
                          "Barents Sea and Norwegian Sea")) %>% 
  mutate(fishingPressureDescription = gsub("Fishing Pressure: " , "", fishingPressureDescription), 
         fishingPressureDescription = gsub("Fishing pressure: " , "", fishingPressureDescription),
         stockSizeDescription = ifelse(stockSizeDescription == "NA", "Stock Size: Relative", stockSizeDescription),
         FmsyDescription = "FMSY",
         stockSizeDescription = gsub("Stock Size: ", "", stockSizeDescription)) %>%
  filter(stockSizeDescription != "B/BMSY" |
           fishingPressureDescription != "F/FMSY") %>%
  group_by(Stock) %>%
  mutate(F_FMSY = F / FMSY,
         SSB_MSYBtrigger = SSB / MSYBtrigger) %>% 
  melt(id.vars = c("AssessmentYear",  "ECOREGION", "GUILD","Stock", "Year"),
       measure.vars = c("F_FMSY", "SSB_MSYBtrigger"),
       variable.name = "METRIC",
       value.name = "VALUE") %>%
  rename(STOCKID = Stock,
         YEAR = Year,
         stockValue = VALUE) %>%
  filter(!is.na(stockValue)) %>% # remove NA VALUE
  group_by(ECOREGION, GUILD, METRIC, YEAR) %>%
  mutate(ecoGuildMean = mean(stockValue)) %>%
  group_by(ECOREGION, METRIC, YEAR) %>%
  mutate(ecoValue = mean(stockValue)) %>%
  group_by(METRIC, GUILD, YEAR) %>%
  mutate(allMean = mean(stockValue))

write.csv(df, file = "~/git/ices-dk/rStockOverview/NordicCouncilRefPoints.csv",
          row.names = FALSE)
# 
```

## Figure 1. ## 

F/F~MSY~ and SSB/MSY B~trigger~ averaged by guild for all ecoregions. [Link to .png](https://github.com/ICES-dk/rStockOverview/blob/master/output/NordicCouncilRefPoints_ALL.png)

```{r echo = FALSE, include = TRUE}
df %>%
  ungroup() %>%
  distinct(GUILD, METRIC, YEAR) %>%
  ggplot(aes(x = YEAR, y = allMean, col = GUILD)) +
  geom_line() +
  geom_hline(yintercept = 1, col = "grey40", linetype = "dashed") +
  # scale_y_continuous(limits = c(0, 3.2)) +
  scale_color_manual(values = c("#7fc97f","#ef3b2c","#662506")) +
  facet_grid(~METRIC) +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 12, angle = 0, hjust = 0),
        strip.background = element_blank(),
        legend.key = element_rect(colour = NA)) +
  labs(x = "Year", y = "Metric")
ggsave(filename = paste0(plotDir, "NordicCouncilRefPoints_ALL.png"), width = 6.5,
       height = 4.5)
```
\pagebreak

## Figure 2. ##  

F/F~MSY~ and SSB/MSY B~trigger~ averaged for each ecoregion and guild. [Link to .png](https://github.com/ICES-dk/rStockOverview/blob/master/output/NordicCouncilRefPoints_ecoregion-guild.png)

```{r echo = FALSE, include = TRUE}

df %>%
  mutate(ECOGUILD = paste0(ECOREGION,", ", GUILD)) %>%
  ungroup() %>%
  # select(ECOGUILD, METRIC, YEAR, ecoGuildMean) %>%
  distinct(ECOGUILD, METRIC, YEAR) %>%
  filter(YEAR >= 1990) %>%
  ggplot(aes(x = YEAR, y = ecoGuildMean, col = METRIC)) +
  geom_line() +
  geom_hline(yintercept = 1, col = "grey40", linetype = "dashed") +
  # scale_y_continuous(limits = c(0, 3.2)) +
  facet_grid(GUILD ~ ECOREGION) +
  scale_color_manual(values = c("#386cb0","#fdb462")) +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 6, angle = 0, hjust = 0),
        axis.text.x = element_text(size = 6, angle = 45),
        strip.background = element_blank(),
        legend.key = element_rect(colour = NA)) +
  labs(x = "Year", y = "Metric")
ggsave(filename = paste0(plotDir, "NordicCouncilRefPoints_ecoregion-guild.png"), width = 6.5,
       height = 4.5)

```
